Sampling-Based Motion Planning for Tracking Evolution of Dynamic Tunnels in Molecular Dynamics Simulations

### 2017-10-31 paper.tex
name of algorithm 6
corrections everywhere

\begin{figure}
\centering
{
\renewcommand{\tabcolsep}{1pt}
\begin{tabular}{cccc}
\includegraphics[width=0.24\textwidth]{fig/blockingb1} &
\includegraphics[width=0.24\textwidth]{fig/blockingb2} & 
\includegraphics[width=0.24\textwidth]{fig/blockingb3} & 
\includegraphics[width=0.24\textwidth]{fig/blockingb4} \\
a & b & c & d \\
\end{tabular}  
}
\caption{\label{fig::blocking1}
    Example of the tree growth without the blocking spheres. 
    The initial configuration $\qinit$ (active site) is represented by the black circle.
    The tree starts to grow inside the protein (a) but it can grow outside as well (b).
    In region denoted by A, the tree grows back to the protein (c).
    Finally, two tunnels are detected (with the exit points $P_1$ and $P_2$), but any tunnel is found in the region $A$ (d).
}
\end{figure}

@@

\begin{algorithm}
%\setstretch{\straa}
{\small
\caption{Tunnel detection in dynamic protein
\label{alg::dynamic}
}
\KwIn{
    starting point $\qinit \in \CF$ in the first frame
}
\KwOut{list of configuration trees $\T_i$, list of exit points $P$, graph $G$ 
}
\hrule
$\T_0$.add($\qinit)$\;
$P = \emptyset$; // list of all exit points \\
$G = $ empty graph\;
\For{$i  = 1 : n$}{
    $(\T_i, \SB)$ = treeTransfer($T_{i-1}, \CF^i$); // Alg.~\ref{alg::transfer}\\
    $(\T_i, P')$ = voidSpaceDetection($\T_{i},\SB)$; //Alg.~\ref{alg::main}\\
    $P=P \cup P'$\;
    $(V,E) =$ voidSpaceMerging($\T_{i}$); // V,E is a graph of merged void space \\
    $C$ = find components in the graph (V,E)\;
    add components $C$ as nodes to graph $G$ \\
    connect exit points $P'$ with their components (nodes) in $G$\;
    connect active site $\qinit \in \CF^i$ with its component (node) in $G$\;
}
}
\end{algorithm}

@@

The number of frames where dynamic tunnels were detected is summarized in Tab.~\ref{tab::dyn}.
In 2,860 frames, no static tunnel was found by CAVER, but  in 1,027 of these cases, a dynamic tunnel was detected and
in other 725 frames, a part of the tunnel (cavity) was detected.
The total number of frames where either end of dynamic tunnel was detected or a cavity that is part of a dynamic tunnel was
1,197.
This shows that the proposed dynamic tunnel detection can bring new information about 1,197 of tunnels, that would be 
discarded in the analysis based solely on static tunnels.
No static tunnel was identified by the proposed method in frames, where no frames were detected by CAVER.
This is not surprising, as there are no active site in these frames so neither static or dynamic tunnel can start in them.
No dynamic neither static tunnel was identified in 1,833 frames and therefore, behavior in these frames  remains unknown.
The results also show that no static tunnel was identified in 669 frames by the proposed method, while CAVER detected static tunnels there.
%The number of frames where the dynamic tunnels were detected is summarized in Tab.~\ref{tab::dyn}.
%In 2,860 frames, no static tunnel was found by CAVER, but in 1,027 frames of this set, a dynamic tunnel was detected.
%Both static and dynamic tunnels were detected in 1,324 frames.
%These results bring new information about the protein behavior. 
%They show that a path connecting the protein surface with the active site is more frequent than the static tunnels suggest.
%The examples of frames without a static tunnel, but containing a dynamic one, are depicted in Fig.~\ref{fig::seq1}.
%
%
%In the case of 3,816 frames where a static tunnel was detected by CAVER, no dynamic tunnel was detected by the proposed method. 
%3,118 frames out of 3,816 were identified as static.
%This indicates that some tunnels were detected as dynamic instead of static.
The proposed method is non-deterministic and it may happen, that some part of the tree does not reach the outer region of the protein and consequently, some exit points are not detected, which decreases the numbef detected static tunnels.
These missed exit points are then reached in the next frame which leads to the detection of the dynamic tunnel.
The number of correctly classified static tunnels can be increased by using higher $\Imax$.
In our experiment, $\Imax=1,200$ was used with a lower value in order to process the whole sequence of 8,000 frames in a reasonable time.

The computation was realized on Intel(R) Xeon(R) CPU E5-2630 v2 @ 2.60GHz with 64GB of RAM using one thread.
As the algorithm holds in the memory only the actual tree (which has less than 10,000 nodes in most of the cases),
the maximally allocated memory during the computation was less than 1GB.
The average computation time to process one frame is 15~s.
In comparison, the computation of the static tunnels using CAVER 3.0 tool takes $\sim$11s per frame.
The processing times for a long sequence of 8,000 frames leads to around 33 hours of computation on a single processor.
Such time is still negligible in the comparison to the preparation of the molecular dynamics simulation.
%Times: 8000 frames, cca 35 hodin = cca 15s/frame. Caver na mem ntb: cca 11s/frame.

The examples of frames without a static tunnel, but with a dynamic one, are depicted in Fig.~\ref{fig::seq1}.
A static tunnel and also a dynamic tunnel was detected in 1324 frames.
These results bring new information about the protein, as they show, that a pathway to the active site exists more
frequently than is described by the classic static tunnels.

>>>>

\begin{figure}
\centering
{
\renewcommand{\tabcolsep}{1pt}
\begin{tabular}{cccc}
\includegraphics[width=0.24\textwidth]{fig/blockingb1} &
\includegraphics[width=0.24\textwidth]{fig/blockingb2} & 
\includegraphics[width=0.24\textwidth]{fig/blockingb3} & 
\includegraphics[width=0.24\textwidth]{fig/blockingb4} \\
a & b & c & d \\
\end{tabular}  
}
\caption{\label{fig::blocking1}
    Example of the tree growth without the blocking spheres. 
    The initial configuration $\qinit$ (active site) is represented by the black circle.
    The tree starts to grow inside the protein (a) but it can grow outside as well (b).
    In region denoted by A, the tree grows back to the protein (c).
    Finally, two tunnels are detected (with the exit points $P_1$ and $P_2$), but no tunnel is found in the region $A$ (d).
}
\end{figure}

@@

\begin{algorithm}
%\setstretch{\straa}
{\small
\caption{Tunnel detection in protein dynamics
\label{alg::dynamic}
}
\KwIn{
    starting point $\qinit \in \CF$ in the first frame
}
\KwOut{list of configuration trees $\T_i$, list of exit points $P$, graph $G$ 
}
\hrule
$\T_0$.add($\qinit)$\;
$P = \emptyset$; // list of all exit points \\
$G = $ empty graph\;
\For{$i  = 1 : n$}{
    $(\T_i, \SB)$ = treeTransfer($T_{i-1}, \CF^i$); // Alg.~\ref{alg::transfer}\\
    $(\T_i, P')$ = voidSpaceDetection($\T_{i},\SB)$; //Alg.~\ref{alg::main}\\
    $P=P \cup P'$\;
    $(V,E) =$ voidSpaceMerging($\T_{i}$); // V,E is a graph of merged void space \\
    $C$ = find components in the graph (V,E)\;
    add components $C$ as nodes to graph $G$ \\
    connect exit points $P'$ with their components (nodes) in $G$\;
    connect active site $\qinit \in \CF^i$ with its component (node) in $G$\;
}
}
\end{algorithm}

@@

The number of frames where dynamic tunnels were detected is summarized in Tab.~\ref{tab::dyn}.
In 2,860 frames, no static tunnel was found by CAVER, but  in 1,027 of these cases, a dynamic tunnel was detected and
in other 725 frames, a part of the tunnel (cavity) was detected.
The total number of frames where either end of a dynamic tunnel was detected or a cavity that is part of a dynamic tunnel was
1,197.
This shows that the proposed dynamic tunnel detection can bring new information about 1,197 of tunnels, that would be 
discarded in the analysis based solely on static tunnels.
No static tunnel was identified by the proposed method in frames, where no static tunnels were detected by CAVER.
As the positions of active site were taken from CAVER, which didn't find suitable place for the large probe in these frames,
also our algorithm could not detect any static tunnels there.
However, our method was able to track evolution of the dynamic tunnels in these frames.
No dynamic neither static tunnel was identified in 1,833 frames and therefore, behavior in these frames  remains unknown.
The results also show that no static tunnel was identified in 669 frames by the proposed method, while CAVER detected static tunnels there.
%The number of frames where the dynamic tunnels were detected is summarized in Tab.~\ref{tab::dyn}.
%In 2,860 frames, no static tunnel was found by CAVER, but in 1,027 frames of this set, a dynamic tunnel was detected.
%Both static and dynamic tunnels were detected in 1,324 frames.
%These results bring new information about the protein behavior. 
%They show that a path connecting the protein surface with the active site is more frequent than the static tunnels suggest.
%The examples of frames without a static tunnel, but containing a dynamic one, are depicted in Fig.~\ref{fig::seq1}.
%
%
%In the case of 3,816 frames where a static tunnel was detected by CAVER, no dynamic tunnel was detected by the proposed method. 
%3,118 frames out of 3,816 were identified as static.
%This indicates that some tunnels were detected as dynamic instead of static.
The proposed method is non-deterministic and it may happen, that some part of the tree does not reach the outer region of the protein and consequently, some exit points are not detected, which decreases the number of detected static tunnels.
These missed exit points are then reached in the next frame which leads to the detection of the dynamic tunnel.
The number of correctly classified static tunnels can be increased by using higher $\Imax$.
In our experiment, $\Imax=1,200$ was used with a lower value in order to process the whole sequence of 8,000 frames in a reasonable time.

The computation was realized on Intel(R) Xeon(R) CPU E5-2630 v2 @ 2.60GHz with 64GB of RAM using one thread.
As the algorithm holds in the memory only the actual tree (which has less than 10,000 nodes in most of the cases),
the maximally allocated memory during the computation was less than 1GB.
The average computation time to process one frame is 15~s.
In comparison, the computation of the static tunnels using CAVER 3.0 tool takes $\sim$11s per frame.
The processing times for a long sequence of 8,000 frames leads to around 33 hours of computation on a single processor.
Such time is still negligible in comparison with the preparation of the molecular dynamics simulation.
%Times: 8000 frames, cca 35 hodin = cca 15s/frame. Caver na mem ntb: cca 11s/frame.

The examples of frames without a static tunnel, but with a dynamic one, are depicted in Fig.~\ref{fig::seq1}.
A static tunnel as well as a dynamic one was detected in 1324 frames.
These results bring new information about the protein, as they show, that a pathway to the active site exists more
frequently than is described by the classic static tunnels.
